---
title: "autostats"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{AutoStats}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(autostats)
```

## plot variable contributions

### multiclass target

Species is a 3-level factor so it will be automatically modelled with a multiclass neural network and a light gbm with multiclass objective function. 

First set define the formula to use for modeling.

```{r}
iris %>% 
  tidy_formula(target = Species) -> species_formula
```


```{r cache=TRUE}
iris %>% 
  auto_variable_contributions(species_formula)
```

```{r cache=TRUE}
iris %>% 
  auto_model_accuracy(species_formula)
```

### binary target 

Linear models uses weighted logistic regression for modeling the coefficients


```{r cache=TRUE}
iris %>% 
  filter(Species != "setosa") %>% 
  auto_variable_contributions(species_formula)
```

For the variable contributions the linear model uses penalized logistic regression provided by glmnet. 

```{r  cache=TRUE}
iris %>% 
  filter(Species != "setosa") -> iris_binary

iris_binary %>% 
  auto_model_accuracy(species_formula)
```

### continuous target

Models are automatically adapted for a continuous target. 

Define the new formula

```{r}
iris %>% 
  tidy_formula(target = Petal.Length) -> petal_formula
```


```{r cache=F}
iris %>% 
  auto_variable_contributions(petal_formula)
```


```{r  cache=TRUE}
iris %>% 
  auto_model_accuracy(petal_formula)
```


